---
title: "Care sites MDD application: precision phenotyping"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE, tidy = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width=6, fig.height=4)
pacman::p_load(data.table, dplyr, foreach, PheWAS, ggplot2, plotly, ggthemr, ggpubr, scales, RColorBrewer, ggrepel, kableExtra, broom, openxlsx)
# setwd("/data/davis_lab/allie/care_sites/scripts")
ggthemr_reset()
```

## Load data
NOTE: the variable names below use "MDD", but the phenotype is actually "depression" (phecode 296.2).
```{r}
## load care site map
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_111524.csv")

## load demographics
dat <- fread("/data/davis_lab/allie/generate_phecode_tables/covariate_files/20230607_sd_wide_covariates_121023.txt")

## load medical home phecodes
phecodes <- readRDS("/data/davis_lab/shared/phenotype_data/biovu/phecode_tables/medical_home_phecode_table_20230607_pull_2_distinct_dates_no_exclusions_121023.Rds")
setDT(phecodes)

## remove phecode columns except for depression
setnames(phecodes, "296.22", "mdd")
setnames(phecodes, "296.2", "dep")

## check that everyone who is a case for 296.22 is also a case for 296.2
stopifnot(nrow(phecodes[dep == 0 & mdd == 1]) == 0)

phe_cols<- names(phecodes)[grepl("^[[:digit:]]", names(phecodes))]
phecodes <- phecodes[,-phe_cols,with=F]
phecodes[, mdd := NULL]
setnames(phecodes,"dep","mdd")

## recode pheno columns
phecodes$mdd <- as.numeric(phecodes$mdd)
phecodes[is.na(mdd), mdd := -9]

## combine phecodes with demographic data
dat <- merge(dat, phecodes, by = "GRID")

head(dat)
```

```{r}
## load SD-wide visit info
## see compile_data.R for details on how this file was created
visit <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")
```

## Depression definitions
* Definition 1: standard depression phecode (296.2)
* Definition 2: 
  * case = has depression phecode (296.2) + ≥1 psych site visit (i.e., has depression and has ever seen a psychiatrist)
  * control = no depression phecode (296.2)
  * excluded = has depression phecode (296.2) but no psych site visit or is N/A for depression phecode
```{r}
psych_sites <- unique(map[SpecialtyLong == "Psychiatry", care_site_id])
psych_grids <- unique(visit[care_site_id %in% psych_sites, GRID])
dat[, psych_cohort := as.numeric(GRID %in% psych_grids)]

dat[, mdd_psych_site := ifelse(mdd==1 & psych_cohort==1, 1,
                                ifelse(mdd==0,0,-9))]

dat[, mdd_pheno := ifelse(mdd_psych_site == 1, "mdd+site", ifelse(mdd == 1, "mdd_no_site", ifelse(mdd == 0, "control", "excluded")))]  
head(dat)                 
```

```{r}
dat %>% 
  group_by(mdd) %>%
  summarise(n = n())
```

## Performance metrics
```{r}
chart_labels <- fread("/data/davis_lab/allie/care_sites/data/mdd_application/tmp.csv") %>%
  select("GRID" = "Subject Numb", "label_long" = "What is your label for this patient?") %>% 
  data.table()

chart_labels$label <- factor(chart_labels$label_long, 
                            levels = c("D No (Not Depression/Not Enough Information)", "D Probable (Lower Evidence)", "D Definite (Higher Evidence)"),
                            labels = c("no_dep", "probable_dep", "definite_dep"))

table(chart_labels$label)
```

```{r}
nrow(chart_labels[GRID %in% dat$GRID])
```
Only 153 of the 200 manually reviewed individuals are in the phecode table.

```{r}
chart_labels <- merge(dat, chart_labels, by = "GRID")

## bin algorithm phenotypes and gold standard labels into cases and non-cases for performance calculations
chart_labels[, case_mdd_code_alg := ifelse(mdd == 1, 1, 0)]
chart_labels[, case_mdd_site_alg := ifelse(mdd_psych_site == 1, 1, 0)]
chart_labels[, case_mdd_chart := ifelse(label == "no_dep", 0, 1)]
```

### Depression codes only
```{r}
tn <- nrow(chart_labels[case_mdd_code_alg == 0 & case_mdd_chart ==  0])
fn <- nrow(chart_labels[case_mdd_code_alg == 0 & case_mdd_chart ==  1])
fp <- nrow(chart_labels[case_mdd_code_alg == 1 & case_mdd_chart ==  0])
tp <- nrow(chart_labels[case_mdd_code_alg == 1 & case_mdd_chart ==  1])

res_mdd_code_alg <- data.table(tp = tp, tn = tn, fp = fp, fn = fn)
res_mdd_code_alg[, ppv := tp/(tp+fp)]
res_mdd_code_alg[, npv := tn/(tn+fn)]
res_mdd_code_alg[, sens := tp/(tp+fn)]
res_mdd_code_alg[, spec := tn/(fp+tn)]
res_mdd_code_alg[, fpr := fp/(fp+tn)]
res_mdd_code_alg[, f1 := 2*tp/(2*tp+fp+fn)]

res_mdd_code_alg
rm(tn, fn, fp, tp)
```

### Depression codes + psych care site visit
```{r}
tn <- nrow(chart_labels[case_mdd_site_alg == 0 & case_mdd_chart ==  0])
fn <- nrow(chart_labels[case_mdd_site_alg == 0 & case_mdd_chart ==  1])
fp <- nrow(chart_labels[case_mdd_site_alg == 1 & case_mdd_chart ==  0])
tp <- nrow(chart_labels[case_mdd_site_alg == 1 & case_mdd_chart ==  1])

res_mdd_site_alg <- data.table(tp = tp, tn = tn, fp = fp, fn = fn)
res_mdd_site_alg[, ppv := tp/(tp+fp)]
res_mdd_site_alg[, npv := tn/(tn+fn)]
res_mdd_site_alg[, sens := tp/(tp+fn)]
res_mdd_site_alg[, spec := tn/(fp+tn)]
res_mdd_site_alg[, fpr := fp/(fp+tn)]
res_mdd_site_alg[, f1 := 2*tp/(2*tp+fp+fn)]

res_mdd_site_alg
rm(tn, fn, fp, tp)
```