---
title: "dep care sites"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table, dplyr, ggplot2, PheWAS, openxlsx, foreach, ggthemr, ggrepel, ggpubr, RColorBrewer, lubridate)
options(scipen=999)
ggthemr_reset()

source("/data/davis_lab/allie/care_sites/scripts/mdd_application/manhattan_bar.R")
```

```{r}
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_111524.csv")
care_site_names <- read.xlsx("/data/davis_lab/allie/care_sites/output/CareSite_VisitDetailCount_UPDATED_111524.xlsx") %>%
  select(care_site_id, care_site_name) %>% 
  data.table() %>% 
  unique()

map <- merge(map, care_site_names, by = "care_site_id")
```

```{r}
phecode_map_full <- data.table(unique(rbind(phecode_map_icd10, phecode_map)))
```

## Load and clean antidepressant data
```{r}
meds <- fread("/data/davis_lab/sealockj/projects/mdd_wbc_long/ad_response/add_antipsychotics/sd_pull_20230607/ad_response/sd_wide_antidepressants_merged_xtable_dbc_from_20230607_pull.csv")
head(meds)
nrow(meds)
```

```{r}
## Select columns of interest 
meds <- meds %>% 
  select(GRID, matched_drug, matched_drug_generic, start_date = drug_exposure_start_date, end_date = drug_exposure_end_date, drug_source_value, visit_occurrence_id, drug_type_concept_name) %>%
  data.table() %>% 
  unique()

nrow(meds)
```

```{r}
## remove if start_date > end_date
meds <- meds[is.na(end_date) | as.Date(start_date) <= as.Date(end_date)]

## remove if start date before 1990 or after 3/31/2023 (date of last ICD or CPT code in the June 2023 SD pull)
meds <- meds[start_date >= as.Date("1990-01-01") & start_date < as.Date("2023-03-31")]

## reformat med column
meds[,matched_drug_generic := tolower(matched_drug_generic)]

nrow(meds)
```

```{r}
sort(table(meds$matched_drug_generic))
```

```{r}
sort(table(meds$drug_type_concept_name))
```

## Select each individual's *earliest* antidepressant record
```{r}
meds[, earliest_med := min(start_date), by = GRID]
meds_earliest <- unique(meds[start_date == earliest_med, .(GRID,matched_drug_generic,start_date,visit_occurrence_id,drug_type_concept_name)])

meds_earliest[,visit_ct := n_distinct(.SD$visit_occurrence_id), by = GRID]
table(meds_earliest$visit_ct)
```

```{r}
sort(table(meds_earliest$matched_drug_generic))
```

```{r}
sort(table(meds_earliest$drug_type_concept_name))
```

```{r}
nrow(meds_earliest[visit_occurrence_id!=0]) / nrow(meds_earliest)
```
Only 34% of med records have a nonzero visit_occurrence_id

```{r}
table(meds_earliest$drug_type_concept_name, meds_earliest$visit_occurrence_id!=0)
```
visit_occurrence_id missingness appears to vary by medication record type

## Map antidepressants to care sites
Map by *date* instead of visit_occurrence_id due to missingness in that column.
```{r, eval = TRUE}
## load all visit data from 2023 SD pull that map to a defined care site (not Unclear or Administrative)
## (generated in compile_data.R)
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")

tot_with_meds <- n_distinct(meds_earliest$GRID) # 667,676 total individuals with an antidepressant record

## map earliest med record to care site by visit_occurrence_id (just to get the exact mapping numbers to report)
meds_visits_by_id <- merge(meds_earliest, vdat, by = c("GRID", "visit_occurrence_id"))

## map earliest med record to care site by DATE
meds_visits <- merge(meds_earliest, vdat, by = "GRID", allow.cartesian = TRUE)
meds_visits <- meds_visits[start_date >= visit_start_date & start_date <= visit_end_date]

n_distinct(meds_visits_by_id$GRID) # 197,687 individuals with earliest med record mapping to a visit with a DEFINED CARE SITE by date (29.6% of total)
n_distinct(meds_visits$GRID) # 354,947 individuals with earliest med record mapping to a visit with a DEFINED CARE SITE by date (53.2% of total)

saveRDS(meds_visits, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date.Rds")
```

```{r}
## load file generated above
meds_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date.Rds")
```


BOOKMARK - EDITING FOR CONSISTENCY WITH mdd_care_sites.Rmd




### Map depression codes to care sites
Try mapping by visit_occurrence_id AND by date.
```{r, eval=FALSE}
## load all visit data from 2023 SD pull that map to a defined care site (not missing, Unclear, or Administrative)
## (generated in compile_data.R)
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")
n_distinct(vdat$GRID) # 3,083,596 total individuals with a visit mapped to a "defined" care site 

## get the total individual count per care site and per specialty
vdat[,cs_n_grid := n_distinct(.SD$GRID), by = care_site_id]
vdat[,spec_n_grid := n_distinct(.SD$GRID), by = SpecialtyLong]
cs_grid_counts <- unique(vdat[,.(care_site_id, care_site_name, cs_n_grid, SpecialtyLong)])
spec_grid_counts <- unique(vdat[,.(SpecialtyLong, spec_n_grid)])
setorder(cs_grid_counts, -cs_n_grid)
setorder(spec_grid_counts, -spec_n_grid)
saveRDS(cs_grid_counts, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_defined_sites_individual_counts_by_site.Rds")
saveRDS(spec_grid_counts, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_defined_sites_individual_counts_by_specialty.Rds")

## map earliest depression icd codes to care sites via visit_occurrence_id
icd_dep_visits_by_id <- merge(icd_dep_earliest, vdat, by = c("GRID","visit_occurrence_id"))
saveRDS(icd_dep_visits_by_id, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_visit_id.Rds")

## map earliest depression icd codes to care sites via date
icd_dep_visits_by_date <- merge(icd_dep_earliest, vdat, by = "GRID", allow.cartesian = TRUE)
icd_dep_visits_by_date <- icd_dep_visits_by_date[entry_date >= visit_start_date & entry_date <= visit_end_date]
saveRDS(icd_dep_visits_by_date, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_date.Rds")

n_distinct(icd_dep_visits_by_id$GRID) # 139082 individuals with earliest depression code mapping to a visit with a DEFINED CARE SITE by date (68.2% of total with depression)
n_distinct(icd_dep_visits_by_date$GRID) # 150206 individuals with earliest dep code mapping to a visit with a DEFINED CARE SITE by date (73.7% of total with depression)
```
Will map by date 1) because there are more individuals and 2) for consistency with the antidepressant analysis.

```{r}
## load file generated above (joined by date)
icd_dep_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_date.Rds")
head(icd_dep_visits)
nrow(icd_dep_visits) # 212406
n_distinct(icd_dep_visits$GRID) # 150206
```

```{r}
## retain unique entries (by GRID, date of ICD code, and care site id)
## there will be duplicates if a care site is mapped to multiple specialties
icd_dep_visits <- unique(icd_dep_visits[,.(GRID,entry_date,care_site_id,care_site_name,SpecialtyLong,cs_n_grid,spec_n_grid)])
nrow(icd_dep_visits) # 186677
n_distinct(icd_dep_visits$GRID) # 150206

nrow(unique(icd_dep_visits[,.(GRID,entry_date,care_site_id)])) # 176580 - number of distinct GRID/care site combos
nrow(unique(icd_dep_visits[,.(GRID,entry_date,care_site_id,care_site_name)])) # 176580 - number of distinct GRID/care site combos (same as above)
nrow(unique(icd_dep_visits[,.(GRID,SpecialtyLong)])) # 181395 - number of distinct GRID/specialty combos
```

Even after restricting to the earliest dep code for each individual and mapping to defined care sites, there are still many individuals with depression codes mapping to multiple different care sites on the same day.
We will keep mutliple entries per individual, but will use the number of unique individuals as the denominator in the care site and specialty frequency calculations.
Otherwise, we would be double-counting individuals with multiple care site visits on the same day. That might bias the results if individuals with dep codes are more likely to have multiple care site visits on the same day, or if certain care sites contribute disproportionately to the total (e.g., radiology or phlebotomy visits, which often co-occur with visits to other sites).

```{r}
paste0("Out of all patients with a depression code, for what % of them does their earliest code map to a visit with a defined specialty? ",
 round((n_distinct(icd_dep_visits$GRID) / n_distinct(icd_dep_earliest$GRID))*100, 2), "%")
```


## Get proportions of individuals with depression codes at each care site and specialty
Normalize by total individual count per care site or specialty.
```{r}
N_dep <- n_distinct(icd_dep_visits$GRID) # 150206 - this will be the denominator for the depression-specific frequencies

icd_dep_visits[,n_dep_per_site := n_distinct(.SD$GRID), by = care_site_id]
icd_dep_visits[,n_dep_per_specialty := n_distinct(.SD$GRID), by = SpecialtyLong]

care_site_summary_dep <- unique(icd_dep_visits[, .(care_site_id, care_site_name, SpecialtyLong, n_dep_per_site, cs_n_grid)])
care_site_summary_dep[,freq_dep_per_site := n_dep_per_site / N_dep]
care_site_summary_dep[,freq_dep_per_site_norm := n_dep_per_site / N_dep / cs_n_grid]

specialty_summary_dep <- unique(icd_dep_visits[, .(SpecialtyLong, n_dep_per_specialty, spec_n_grid)])
specialty_summary_dep[,freq_dep_per_specialty := n_dep_per_specialty / N_dep]
specialty_summary_dep[,freq_dep_per_specialty_norm := n_dep_per_specialty / N_dep / spec_n_grid]
```

## Plots
```{r}
## reformat data frame for plotting
caresite_plot_df <- care_site_summary_dep %>% 
  filter(SpecialtyLong != "Phlebotomy" & SpecialtyLong != "Radiology"  & SpecialtyLong != "Research") %>%
  filter(n_dep_per_site >=100) %>%
  select(care_site_id, care_site_name, 
    specialty = SpecialtyLong, n_phe_per_site = n_dep_per_site, 
    freq_phe_per_site = freq_dep_per_site, freq_phe_per_site_norm = freq_dep_per_site_norm) %>%
  data.table()

p1 <- caresite_manhattan_bar(caresite_plot_df, points = T, flip = F) +
  labs(x = "", y = "Frequency of depression codes per care site\n")

## create a second version of the plot, normalizing by total individual count per care site
caresite_plot_df_norm <- caresite_plot_df %>%
  select(-freq_phe_per_site) %>% 
  rename(freq_phe_per_site = freq_phe_per_site_norm) %>% 
  mutate(freq_phe_per_site = freq_phe_per_site * 10000) %>%
  data.table()

p2 <- caresite_manhattan_bar(caresite_plot_df_norm, points = T, flip = T) +
  labs(x = "", y = "Frequency of depression codes per care site\n(normalized by per-site individual count)")

p <- ggarrange(p1, p2, ncol=1,nrow=2)

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/earliest_dep_site_counts_bar_mapped_by_date_miami_points.pdf", width = 11, height = 8.5)
print(p)
dev.off()
```

```{r}
print(p)
```