---
title: "Antidepressant care sites"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table, dplyr, ggplot2, PheWAS, openxlsx, foreach, ggthemr, ggrepel, RColorBrewer, lubridate, kableExtra)
options(scipen=999)

ggthemr_reset()
```

TODO:
* Update denominator to be consistent with mdd_care_sites.Rmd
* Generally update for consistency with mdd_care_sites.Rmd

## Load data
```{r}
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_111524.csv")
care_site_names <- read.xlsx("/data/davis_lab/allie/care_sites/output/CareSite_VisitDetailCount_UPDATED_111524.xlsx") %>%
  select(care_site_id, care_site_name) %>% 
  data.table() %>% 
  unique()

map <- merge(map, care_site_names, by = "care_site_id")
```

### Load and clean antidepressant data
```{r}
meds <- fread("/data/davis_lab/sealockj/projects/mdd_wbc_long/ad_response/add_antipsychotics/sd_pull_20230607/ad_response/sd_wide_antidepressants_merged_xtable_dbc_from_20230607_pull.csv")
nrow(meds)
```

```{r}
## Select columns of interest 
meds <- meds %>% 
  select(GRID, matched_drug, matched_drug_generic, start_date = drug_exposure_start_date, end_date = drug_exposure_end_date, drug_source_value, visit_occurrence_id, drug_type_concept_name) %>%
  data.table() %>% 
  unique()

nrow(meds)
```

```{r}
## remove if start_date > end_date
meds <- meds[is.na(end_date) | as.Date(start_date) <= as.Date(end_date)]

## remove if start date before 1990 or after 3/31/2023 (date of last ICD or CPT code in the June 2023 SD pull)
meds <- meds[start_date >= as.Date("1990-01-01") & start_date < as.Date("2023-03-31")]

## reformat med column
meds[,matched_drug_generic := tolower(matched_drug_generic)]

nrow(meds)
```

```{r}
sort(table(meds$matched_drug_generic))
```

```{r}
sort(table(meds$drug_type_concept_name))
```

## Select each individual's *earliest* antidepressant record
```{r}
meds[, earliest_med := min(start_date), by = GRID]
meds_earliest <- unique(meds[start_date == earliest_med, .(GRID,matched_drug_generic,start_date,visit_occurrence_id,drug_type_concept_name)])

meds_earliest[,visit_ct := n_distinct(.SD$visit_occurrence_id), by = GRID]
table(meds_earliest$visit_ct)
```

```{r}
sort(table(meds_earliest$matched_drug_generic))
```

```{r}
sort(table(meds_earliest$drug_type_concept_name))
```

```{r}
nrow(meds_earliest[visit_occurrence_id!=0]) / nrow(meds_earliest)
```
Only 34% of med records have a nonzero visit_occurrence_id

```{r}
table(meds_earliest$drug_type_concept_name, meds_earliest$visit_occurrence_id!=0)
```
Missingness varies by medication record type.

## Map antidepressants to care sites
Map by *date* instead of visit_occurrence_id due to missingness in that column.
```{r, eval = FALSE}
## load all visit data from 2023 SD pull that map to a defined care site (not Unclear or Administrative)
## (generated in compile_data.R)
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")

tot_with_meds <- n_distinct(meds_earliest$GRID) # 667,676 total individuals with an antidepressant record

## map earliest med record to care site by visit_occurrence_id (just to get the exact mapping numbers to report)
meds_visits_by_id <- merge(meds_earliest, vdat, by = c("GRID", "visit_occurrence_id"))

## map earliest med record to care site by DATE
meds_visits <- merge(meds_earliest, vdat, by = "GRID", allow.cartesian = TRUE)
meds_visits <- meds_visits[start_date >= visit_start_date & start_date <= visit_end_date]

n_distinct(meds_visits_by_id$GRID) # 197,687 individuals with earliest med record mapping to a visit with a DEFINED CARE SITE by date (29.6% of total)
n_distinct(meds_visits$GRID) # 354,947 individuals with earliest med record mapping to a visit with a DEFINED CARE SITE by date (53.2% of total)

saveRDS(meds_visits, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date.Rds")
```

```{r}
## load file generated above
meds_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date.Rds")
```

### summarize missingness by drug_type_concept_name
```{r}
meds_summary <- meds_earliest %>% 
  group_by(drug_type_concept_name) %>%
  summarize(n_with_med = n_distinct(GRID), median_start_year = median(year(start_date)), max_start_year = max(year(start_date)))

meds_visits_summary <- meds_visits %>% 
  group_by(drug_type_concept_name) %>%
  summarize(n_with_med_and_visit = n_distinct(GRID), median_start_year = median(year(start_date)), max_start_year = max(year(start_date)))

overall_summary <- merge(meds_summary, meds_visits_summary, by = "drug_type_concept_name")
overall_summary$prop_with_visit <- overall_summary$n_with_med_and_visit / overall_summary$n_with_med

overall_summary %>% 
  arrange(desc(prop_with_visit)) %>% 
  kable() %>%
  kable_styling()
```
When mapping meds to visits (with defined care sites) by date, within the Prescription written and Medication list entry groups, we're seeing about the same missingness rates as with the ICD codes.

```{r, eval = FALSE}
## restrict to prescriptions / physician administered meds
meds_visits <- meds_visits[drug_type_concept_name != "Medication list entry" & drug_type_concept_name != "Patient Self-Reported Medication"]
```

```{r}
meds_visits <- unique(meds_visits[,.(GRID,matched_drug_generic,start_date,care_site_id,care_site_name,visit_type,SpecialtyLong)])

## assign unique visit id based on GRID/med start_date/care_site_id combo
## this is so that we have an accurate denominator for the proportion of antidepressant visits mapping to each care site
## this will prevent double counting of visits mapping to care sites which map to multiple specialties
meds_visits[,my_visit_id := as.integer(factor(paste0(GRID, start_date, care_site_id, care_site_name)))]

meds_visits[, visit_ct := n_distinct(.SD$my_visit_id), by = GRID]
table(meds_visits$visit_ct)
```
Even after restricting to the earliest antidepressant for each individual and mapping to defined care sites, there are still some individuals with med records mapping to multiple visits/care sites. 
We will keep mutliple entries per individual for now.

## Get proportions of earliest antidepressant visits that are documented in each care site
```{r}
meds_visits[,n_per_site := n_distinct(.SD$my_visit_id), by = care_site_id]
meds_visits[,n_per_specialty := n_distinct(.SD$my_visit_id), by = SpecialtyLong]

n_tot_med <- n_distinct(meds_visits$my_visit_id) # this is the total number of med visits mapping to a care site with a defined specialty
n_tot_med
```

```{r}
care_site_summary_med <- unique(meds_visits[, .(care_site_id, care_site_name, SpecialtyLong, n_per_site)])
care_site_summary_med[,freq_per_site := n_per_site / n_tot_med]

specialty_summary_med <- unique(meds_visits[, .(SpecialtyLong, n_per_specialty)])
specialty_summary_med[,freq_per_specialty := n_per_specialty / n_tot_med]

setnames(care_site_summary_med, "n_per_site", "n_med")
setnames(care_site_summary_med, "freq_per_site", "freq_med")
setnames(specialty_summary_med, "n_per_specialty", "n_med")
setnames(specialty_summary_med, "freq_per_specialty", "freq_med")
```

```{r}
care_site_summary <- care_site_summary_med %>% 
  select(care_site_id, care_site_name, specialty = SpecialtyLong,  med_in = n_med, freq_med) %>%
  data.table()

care_site_summary[,med_out := n_tot_med - med_in] # med_out = number of med visits NOT mapping to this care site

specialty_summary <- specialty_summary_med %>% 
  select(specialty = SpecialtyLong, med_in = n_med, freq_med) %>%
  data.table()

specialty_summary[,med_out := n_tot_med - med_in] # med_out = number of med visits NOT mapping to this specialty
```

## Plots
```{r}
## restrict to care sites with >=500 antidepressant visits
caresite_plot_df <- care_site_summary %>% 
  select(care_site_id, care_site_name, MappedSpecialty = specialty, n_med_visits = med_in, perc_of_med_visits = freq_med) %>%
  filter(n_med_visits >=500) %>%
  group_by(MappedSpecialty) %>%
  mutate(N = n()) %>%  ## number of data points per specialty, for plotting purposes
  data.table()

caresite_plot_df[, N := N+2]

Specialties_3group <- fread("/data/davis_lab/allie/care_sites/output/Figures/Specialties_NCareSites_Label_ANNOTATED.csv") %>% 
  filter(SpecialtyLong %in% caresite_plot_df$MappedSpecialty) %>% 
  select(MappedSpecialty = SpecialtyLong, Group) %>%
  data.table()

Specialties_3group <- merge(Specialties_3group, unique(caresite_plot_df[,.(MappedSpecialty, N)]), by = "MappedSpecialty")
Specialties_3group$Group <- factor(Specialties_3group$Group, levels = c("Medical", "Surgical", "Other"))

# Create group to alternate light/dark shades
group_colors <- c("Medical_light" = "#a6cee3", "Medical_dark" = "#1f78b4",
                    "Surgical_light" = "#b2df8a", "Surgical_dark" = "#33a02c",
                    "Other_light" = "#fb9a99", "Other_dark" = "#e31a1c")

# Create a base value for each specialty
specialty_base <- Specialties_3group %>%
  arrange(Group, MappedSpecialty) %>%
  distinct(MappedSpecialty, N) %>%
  mutate(base_value = cumsum(N) - N)

# Order the care sites
caresite_plot_df <- merge(caresite_plot_df, Specialties_3group[,.(MappedSpecialty, Group)], by = "MappedSpecialty")
caresite_plot_df <- merge(caresite_plot_df, specialty_base[,.(MappedSpecialty, base_value)], by = "MappedSpecialty")
setDT(caresite_plot_df)
setorder(caresite_plot_df, MappedSpecialty, care_site_name)
caresite_plot_df[, id := seq_len(.N), by = MappedSpecialty]
caresite_plot_df[, id := rowid(MappedSpecialty)]
caresite_plot_df[,RowNumber := base_value + id]

caresite_plot_df <- caresite_plot_df %>%
  group_by(MappedSpecialty) %>%
  mutate(Group_Shade = ifelse(cur_group_id() %% 2 == 0, paste(Group, "light", sep = "_"), paste(Group, "dark", sep = "_"))) %>%
  mutate(Specialty_midpoint = median(RowNumber)) %>% 
  ungroup() %>% 
  data.table()

midpoint_labels <- caresite_plot_df %>%
  group_by(MappedSpecialty) %>%
  summarize(midpoint = median(RowNumber)) %>%
  ungroup()

if (FALSE) {
  N_spec <- n_distinct(caresite_plot_df$MappedSpecialty)
  pal <- rev(rep(brewer.pal(n = 8, name = "Set2"),6)[1:N_spec])
  spec_order <- unique(caresite_plot_df[,.(MappedSpecialty)])
  spec_order$Color <- pal
  caresite_plot_df <- merge(caresite_plot_df, spec_order, by = "MappedSpecialty")
}

p <- ggplot(caresite_plot_df, aes(x = RowNumber, y = perc_of_med_visits, fill = Group_Shade)) +
  geom_bar(stat = "identity", color = "black") +
  geom_label_repel(data = caresite_plot_df %>% filter(perc_of_med_visits >= 0.01) %>% arrange(desc(n_med_visits)) %>% group_by(MappedSpecialty) %>% slice(1:5), 
                  aes(label = care_site_name),
                  size = 3,
                  box.padding = 0.5, point.padding = 0.5,
                  vjust = 0.5, hjust = 0.5, max.overlaps = 20) +
  scale_x_continuous(breaks = midpoint_labels$midpoint, labels = midpoint_labels$MappedSpecialty) +
  scale_fill_manual(values = group_colors) +
  # scale_fill_identity()  +
  theme_minimal(base_family = "Helvetica") +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 8), 
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"),
        legend.position = "none", 
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) + 
  labs(x = "Care site", y = "Proportion of index antidepressant visits")

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/earliest_antidepressant_site_counts_bar.pdf", width = 10, height = 7)
print(p)
dev.off()
```

```{r}
print(p)
```