---
title: "dep care sites"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table, dplyr, ggplot2, PheWAS, openxlsx, foreach, ggthemr, ggrepel, ggpubr, RColorBrewer, lubridate)
options(scipen=999)
ggthemr_reset()

source("/data/davis_lab/allie/care_sites/scripts/mdd_application/manhattan_bar.R")
```

```{r}
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_AML_010425.csv")
care_site_names <- read.xlsx("/data/davis_lab/allie/care_sites/output/CareSite_VisitDetailCount_UPDATED_AML_010425.xlsx") %>%
  select(care_site_id, care_site_name) %>% 
  data.table() %>% 
  unique()

map <- merge(map, care_site_names, by = "care_site_id")
```

## Load and clean antidepressant data
```{r}
meds <- fread("/data/davis_lab/sealockj/projects/mdd_wbc_long/ad_response/add_antipsychotics/sd_pull_20230607/ad_response/sd_wide_antidepressants_merged_xtable_dbc_from_20230607_pull.csv")
head(meds)
nrow(meds)
```

```{r}
## Select columns of interest 
meds <- meds %>% 
  select(GRID, matched_drug, matched_drug_generic, start_date = drug_exposure_start_date, end_date = drug_exposure_end_date, drug_source_value, visit_occurrence_id, drug_type_concept_name) %>%
  data.table() %>% 
  unique()

nrow(meds)
```

```{r}
## remove if start_date > end_date
meds <- meds[is.na(end_date) | as.Date(start_date) <= as.Date(end_date)]

## remove if start date before 1998 (beginning of StarPanel) or after 3/31/2023 (date of last ICD or CPT code in the June 2023 SD pull)
meds <- meds[start_date >= as.Date("1998-01-01") & start_date < as.Date("2023-03-31")]

## reformat med column
meds[,matched_drug_generic := tolower(matched_drug_generic)]

nrow(meds)
```

```{r}
sort(table(meds$matched_drug_generic))
```

```{r}
sort(table(meds$drug_type_concept_name))
```

### Remove self-reported medications and med list entries.
```{r}
meds <- meds[drug_type_concept_name != "Medication list entry" & drug_type_concept_name != "Patient Self-Reported Medication"]
```

```{r}
sort(table(meds$drug_type_concept_name))
```

## Select each patient's *earliest* antidepressant record
```{r}
meds[, earliest_code := min(start_date), by = GRID]
meds_earliest <- unique(meds[start_date == earliest_code, .(GRID, start_date, visit_occurrence_id, matched_drug_generic, drug_type_concept_name)])

## collapse multiple medications per individual
meds_earliest[, meds := paste0(unique(matched_drug_generic), collapse = ", "), by = GRID]
meds_earliest <- unique(meds_earliest[,.(GRID,start_date,visit_occurrence_id,meds, drug_type_concept_name)])

n_distinct(meds_earliest$GRID) # 325698 total individuals with an antidepressant prescription or administration
nrow(meds_earliest) # 458063 total entries
```

```{r}
meds_earliest %>% 
  group_by(meds) %>% 
  summarise(n = n()) %>%
  arrange(-n) %>%
  head(20)
```

```{r}
nrow(meds_earliest[visit_occurrence_id!=0]) / nrow(meds_earliest)
```
65.1% of med records have a nonzero visit_occurrence_id

```{r}
meds_earliest %>% 
  group_by(drug_type_concept_name) %>%
  summarise(prop_missing = sum(visit_occurrence_id==0) / n())
```
visit_occurrence_id missingness appears to vary by medication record type

## Map antidepressants to care sites
Map by *date* instead of visit_occurrence_id due to missingness in that column.
```{r, eval = TRUE}
## load all visit data from 2023 SD pull that map to a defined care site (not Unclear or Administrative)
## (generated in compile_data.R)
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites_010525.Rds")
n_distinct(vdat$GRID) # 2959903 total individuals with a visit mapped to a "defined" care site 

## get the total individual count per care site and per specialty
vdat[,cs_n_grid := n_distinct(.SD$GRID), by = care_site_id]
vdat[,spec_n_grid := n_distinct(.SD$GRID), by = SpecialtyLong]

## map earliest antidepressant icd codes to care sites via visit_occurrence_id
med_visits_by_id <- merge(meds_earliest, vdat, by = c("GRID","visit_occurrence_id"))
saveRDS(med_visits_by_id, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_visit_id_010525.Rds")

## map earliest antidepressant icd codes to care sites via date
med_visits_by_date <- merge(meds_earliest, vdat, by = "GRID", allow.cartesian = TRUE)
med_visits_by_date <- med_visits_by_date[start_date >= visit_start_date & start_date <= visit_end_date]
saveRDS(med_visits_by_date, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date_010525.Rds")

n_distinct(med_visits_by_id$GRID) # 129973 individuals with earliest antidepressant mapping to a visit with a DEFINED CARE SITE by date (39.9% of total with an antidepressant)
n_distinct(med_visits_by_date$GRID) # 215510 individuals with earliest antidepressant mapping to a visit with a DEFINED CARE SITE by date (66.2% of total with an antidepressant)
```

```{r}
## load file generated above (joined by date)
med_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_med_joined_by_date_010525.Rds")
head(med_visits)
nrow(med_visits) # 393300
n_distinct(med_visits$GRID) # 215510
```

```{r}
## retain unique entries (by GRID, med start date, care site, and specialty)
## there will be duplicates if a care site is mapped to multiple specialties
med_visits <- unique(med_visits[,.(GRID,start_date,care_site_id,care_site_name,SpecialtyLong,cs_n_grid,spec_n_grid)])
nrow(med_visits) #  261816
n_distinct(med_visits$GRID) # 215510

nrow(unique(med_visits[,.(GRID,start_date,care_site_id)])) # 245657 - number of distinct GRID/care site combos
nrow(unique(med_visits[,.(GRID,start_date,care_site_id,care_site_name)])) # 245657 - number of distinct GRID/care site combos (same as above)
nrow(unique(med_visits[,.(GRID,SpecialtyLong)])) # 254192 - number of distinct GRID/specialty combos
```

Even after restricting to the earliest antidepressant for each individual and mapping to defined care sites, there are still many individuals with meds mapping to multiple different care sites on the same day.
We will keep mutliple entries per individual, but will use the number of unique individuals as the denominator in the care site and specialty frequency calculations.
Otherwise, we would be double-counting individuals with multiple care site visits on the same day. That might bias the results if individuals with antidepressants are more likely to have multiple care site visits on the same day, or if certain care sites contribute disproportionately to the total (e.g., radiology or phlebotomy visits, which often co-occur with visits to other sites).

```{r}
paste0("Out of all patients with a med, for what % of them does their earliest code map to a visit with a defined specialty? ",
 round((n_distinct(med_visits$GRID) / n_distinct(meds_earliest$GRID))*100, 2), "%")
```

## Get proportions of individuals with antidepressants at each care site and specialty
Normalize by total individual count per care site or specialty.
```{r}
n_med <- n_distinct(med_visits$GRID) # 215510 - this will be the denominator for the med-specific frequencies

med_visits[,n_med_per_site := n_distinct(.SD$GRID), by = care_site_id]
med_visits[,n_med_per_specialty := n_distinct(.SD$GRID), by = SpecialtyLong]

care_site_summary_meds <- unique(med_visits[, .(care_site_id, care_site_name, SpecialtyLong, n_med_per_site, cs_n_grid)])
care_site_summary_meds[,freq_med_per_site := n_med_per_site / n_med]
care_site_summary_meds[,freq_med_per_site_norm := n_med_per_site / n_med / cs_n_grid]

specialty_summary_meds <- unique(med_visits[, .(SpecialtyLong, n_med_per_specialty, spec_n_grid)])
specialty_summary_meds[,freq_med_per_specialty := n_med_per_specialty / n_med]
specialty_summary_meds[,freq_med_per_specialty_norm := n_med_per_specialty / n_med / spec_n_grid]
```

## Make supplemental tables
```{r}
care_site_supp <- copy(care_site_summary_meds)
names(care_site_supp) <- c("care_site_id", "care_site_name", "specialty", "n_antidepressant_patients_per_site", "n_total_patients_per_site", "prop_antidepressant_patients_per_site","prop_antidepressant_patients_per_site_normalized")
setorder(care_site_supp, -n_antidepressant_patients_per_site)

specialty_supp <- copy(specialty_summary_meds)
names(specialty_supp) <- c("specialty", "n_antidepressant_patients_per_specialty", "n_total_patients_per_specialty", "prop_antidepressant_patients_per_specialty","prop_antidepressant_patients_per_specialty_normalized")
setorder(specialty_supp, -n_antidepressant_patients_per_specialty)

## write to excel
wb <- createWorkbook()
addWorksheet(wb, "Care Site Summary")
writeData(wb, "Care Site Summary", care_site_supp)
addWorksheet(wb, "Specialty Summary")
writeData(wb, "Specialty Summary", specialty_supp)
saveWorkbook(wb, "/data/davis_lab/allie/care_sites/output/mdd_application/SupplementalTable_antidepressant_care_site_specialty_summary_010525.xlsx", overwrite = TRUE)
```

```{r}
# final dataset- get GRID and med date
med_date <- unique(med_visits[,.(GRID, start_date)])

# re-extract med records
med_list <- merge(meds, med_date, by = c("GRID", "start_date")) 
n_distinct(med_list$GRID) # 215510

med_list <- unique(med_list[,.(GRID, generic_name = toupper(matched_drug_generic))])

med_map <- unique(meds[,.(generic_name = toupper(matched_drug_generic), brand_name = matched_drug)])
med_map <- med_map[brand_name != generic_name]
med_map[, all_brand_names := paste0(unique(brand_name), collapse = ", "), by = generic_name]
med_map <- unique(med_map[,.(generic_name, all_brand_names)])

med_list <- merge(med_list, med_map, by = "generic_name", allow.cartesian = TRUE)

med_list_tbl <- med_list %>% 
  group_by(generic_name, all_brand_names) %>% 
  summarise(n = n()) %>%
  arrange(-n) %>% 
  as.data.frame()

## write to excel
wb <- createWorkbook()
addWorksheet(wb, "Antidepressant List")
writeData(wb, "Antidepressant List", med_list_tbl)
saveWorkbook(wb, "/data/davis_lab/allie/care_sites/output/mdd_application/SupplementalTable_antidepressant_list_010525.xlsx", overwrite = TRUE)
```

## Plots
```{r}
## reformat data frame for plotting
caresite_plot_df <- care_site_summary_meds %>% 
  filter(SpecialtyLong != "Phlebotomy" & SpecialtyLong != "Radiology"  & SpecialtyLong != "Research") %>%
  filter(n_med_per_site >=100) %>%
  select(care_site_id, care_site_name, 
    specialty = SpecialtyLong, n_phe_per_site = n_med_per_site, 
    freq_phe_per_site = freq_med_per_site, freq_phe_per_site_norm = freq_med_per_site_norm) %>%
  data.table()

p1 <- caresite_manhattan_bar(caresite_plot_df, points = T, flip = F) +
  labs(x = "", y = "Proportion of prescriptions per care site\n") +
  scale_y_continuous(breaks = seq(0,0.03,0.005))

## create a second version of the plot, normalizing by total individual count per care site
caresite_plot_df_norm <- caresite_plot_df %>%
  select(-freq_phe_per_site) %>% 
  rename(freq_phe_per_site = freq_phe_per_site_norm) %>% 
  mutate(freq_phe_per_site = freq_phe_per_site * 10000) %>%
  data.table()

p2 <- caresite_manhattan_bar(caresite_plot_df_norm, points = T, flip = T) +
  labs(x = "", y = "Proportion of prescriptions per care site\n(normalized by per-site individual count)") +
  scale_y_reverse(breaks = rev(seq(0,0.03,0.005)), limits = c(0.025, 0))

p <- ggarrange(p1, p2, ncol=1,nrow=2)

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/earliest_med_site_counts_mapped_by_date_miami_points_010525.pdf", width = 11, height = 8.5)
print(p)
dev.off()
```

```{r}
print(p)
```
