---
title: "MDD care sites"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table, dplyr, ggplot2, PheWAS, openxlsx, foreach, ggthemr, ggrepel, RColorBrewer, lubridate)
options(scipen=999)

ggthemr_reset()
```

## Load data
```{r}
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_092424.csv")
care_site_names <- read.xlsx("/data/davis_lab/allie/care_sites/output/CareSite_VisitDetailCount_UPDATED_092424.xlsx") %>%
  select(care_site_id, care_site_name) %>% 
  data.table() %>% 
  unique()

map <- merge(map, care_site_names, by = "care_site_id")
```

```{r}
phecode_map_full <- data.table(unique(rbind(phecode_map_icd10, phecode_map)))
```

```{r load_icd, eval=FALSE}
icd <- readRDS("/data/davis_lab/allie/commonly_used/20230607_sd_pull_x_codes_icd.Rds")
dep_codes <- phecode_map_full[phecode %in% c("296.2", "296.22")]

icd_dep <- unique(icd[concept_code %in% dep_codes$code, .(GRID, entry_date, entry_datetime, 
                                                          vocabulary_id, concept_code, concept_name, 
                                                          visit_occurrence_id, item_position, age_at_event)])

saveRDS(icd_dep, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression.Rds")
```

```{r}
icd_dep <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression.Rds") %>% 
  select(GRID, entry_date, entry_datetime, vocabulary_id, concept_code, concept_name, visit_occurrence_id, age_at_event) %>% 
  data.table() %>% 
  unique()

icd_dep$entry_datetime <- as_datetime(icd_dep$entry_datetime)

table(icd_dep$vocabulary_id)
icd_dep$vocabulary_id <- gsub("CM", "", icd_dep$vocabulary_id)
table(icd_dep$vocabulary_id)

## restrict to MDD only (296.22)
mdd_codes <- phecode_map_full[phecode == "296.22"]
setnames(mdd_codes, "code", "concept_code")
table(mdd_codes$vocabulary_id)
mdd_codes$vocabulary_id <- gsub("CM", "", mdd_codes$vocabulary_id)
mdd_codes <- unique(mdd_codes)
table(mdd_codes$vocabulary_id)

icd_mdd <- merge(mdd_codes, icd_dep, by = c("concept_code", "vocabulary_id"))
icd_mdd[,code_index := .I]
```

## Get proportion of each care site among all visits mapping to sites with defined specialties (baseline frequencies)
```{r, eval = F}
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")

vdat[,n_per_site := length(unique(.SD$visit_occurrence_id)), by = care_site_id]
vdat[,n_per_specialty := length(unique(.SD$visit_occurrence_id)), by = SpecialtyLong]

n_tot <- length(unique(vdat$visit_occurrence_id)) # 46422711 (this is the total number of visits mapping to a care site with a defined specialty)

care_site_summary <- unique(vdat[, .(care_site_id, care_site_name, SpecialtyLong, n_per_site)])
care_site_summary[,freq_per_site := n_per_site / n_tot]

specialty_summary <- unique(vdat[, .(SpecialtyLong, n_per_specialty)])
specialty_summary[,freq_per_specialty := n_per_specialty / n_tot]

saveRDS(care_site_summary, "/data/davis_lab/allie/care_sites/data/mdd_application/care_site_frequencies.Rds")
saveRDS(specialty_summary, "/data/davis_lab/allie/care_sites/data/mdd_application/specialty_frequencies.Rds")
```

## Select each individual's *earliest* MDD code
```{r}
icd_mdd[, earliest_code := min(entry_datetime), by = GRID]
icd_mdd_earliest <- unique(icd_mdd[entry_datetime == earliest_code, .(GRID,entry_date,entry_datetime,visit_occurrence_id)])

icd_mdd_earliest[,visit_ct := .N, by = GRID]
table(icd_mdd_earliest$n_icd)
```

### Map MDD codes to care sites by visit_occurrence_id
```{r, eval = F}
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")
icd_mdd_visits <- merge(icd_mdd_earliest, vdat, by = c("GRID","visit_occurrence_id"))
saveRDS(icd_mdd_visits, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_mdd_icd_joined_by_visit_id.Rds")
```

```{r}
icd_mdd_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_mdd_icd_joined_by_visit_id.Rds")

icd_mdd_visits[, visit_ct := length(unique(.SD$visit_occurrence_id)), by = GRID]
icd_mdd_visits[, visit_entry_ct := .N, by = GRID] # accounts for multiple care sites per visit
table(icd_mdd_visits$visit_ct)
table(icd_mdd_visits$visit_entry_ct)
```

Even after restricting to the earliest MDD code for each individual and mapping to defined care sites, there are still some individuals with multiple MDD codes. 
We will keep mutliple entries per individual for now.
```{r}
n_distinct(icd_mdd_earliest$GRID)
n_distinct(icd_mdd_visits$GRID)

paste0("Proportion of individuals with at least one (earliest) MDD code mapping to a visit with a defined specialty: ",
 round(n_distinct(icd_mdd_visits$GRID) / n_distinct(icd_mdd_earliest$GRID), 3))
```

## Get proportion of each care site among all *earliest* MDD visits (i.e. MDD visits are first documented in care site X, X% of the time)
```{r}
icd_mdd_visits[,n_per_site := length(unique(.SD$visit_occurrence_id)), by = care_site_id]
icd_mdd_visits[,n_per_specialty := length(unique(.SD$visit_occurrence_id)), by = SpecialtyLong]

n_tot_mdd <- length(unique(icd_mdd_visits$visit_occurrence_id)) # this is the total number of MDD visits mapping to a care site with a defined specialty

care_site_summary_mdd <- unique(icd_mdd_visits[, .(care_site_id, care_site_name, SpecialtyLong, n_per_site)])
care_site_summary_mdd[,freq_per_site := n_per_site / n_tot_mdd]

specialty_summary_mdd <- unique(icd_mdd_visits[, .(SpecialtyLong, n_per_specialty)])
specialty_summary_mdd[,freq_per_specialty := n_per_specialty / n_tot_mdd]

setnames(care_site_summary_mdd, "n_per_site", "n_mdd")
setnames(care_site_summary_mdd, "freq_per_site", "freq_mdd")
setnames(specialty_summary_mdd, "n_per_specialty", "n_mdd")
setnames(specialty_summary_mdd, "freq_per_specialty", "freq_mdd")
```

### Combine with baseline frequencies
```{r}
care_site_summary <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/care_site_frequencies.Rds") %>%
  rename(n_tot = n_per_site, freq_tot = freq_per_site) %>%
  data.table()

specialty_summary <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/specialty_frequencies.Rds") %>% 
  rename(n_tot = n_per_specialty, freq_tot = freq_per_specialty) %>%
  data.table()

N <- 46422711 # (this is the total number of visits mapping to a care site with a defined specialty)
N_mdd <- n_tot_mdd # 468107 (this is the total number of MDD visits mapping to a care site with a defined specialty)

care_site_summary <- merge(care_site_summary, care_site_summary_mdd, by = c("care_site_id", "care_site_name", "SpecialtyLong"))
specialty_summary <- merge(specialty_summary, specialty_summary_mdd, by = "SpecialtyLong")

care_site_summary <- care_site_summary %>% 
  select(care_site_id, care_site_name, specialty = SpecialtyLong, tot_in = n_tot, mdd_in = n_mdd, freq_tot, freq_mdd) %>%
  data.table()

care_site_summary[,tot_out := N - tot_in]
care_site_summary[,mdd_out := N_mdd - mdd_in]

specialty_summary <- specialty_summary %>% 
  select(specialty = SpecialtyLong, tot_in = n_tot, mdd_in = n_mdd, freq_tot, freq_mdd) %>%
  data.table()

specialty_summary[,tot_out := N - tot_in]
specialty_summary[,mdd_out := N_mdd - mdd_in]
```

### Chi-square tests
```{r, eval=FALSE}
care_site_summary <- foreach(i = 1:nrow(care_site_summary), .combine = rbind) %do% { 
  row <- care_site_summary[i,]
  mat <- matrix(c(row$tot_in, row$tot_out, row$mdd_in, row$mdd_out), nrow = 2)
  test <- chisq.test(mat)
  row$chisq.stat <- test$statistic
  row$chisq.pval <- test$p.value
  row
}

specialty_summary <- foreach(i = 1:nrow(specialty_summary), .combine = rbind) %do% { 
  row <- specialty_summary[i,]
  mat <- matrix(c(row$tot_in, row$tot_out, row$mdd_in, row$mdd_out), nrow = 2)
  test <- chisq.test(mat)
  row$chisq.stat <- test$statistic
  row$chisq.pval <- test$p.value
  row
}
```

## Plots
```{r}
## restrict to care sites with >=100 MDD visits
caresite_plot_df <- care_site_summary %>% 
  select(care_site_id, care_site_name, MappedSpecialty = specialty, n_mdd_visits = mdd_in, perc_of_mdd_visits = freq_mdd) %>%
  filter(n_mdd_visits >=100) %>%
  group_by(MappedSpecialty) %>%
  mutate(N = n()) %>% 
  data.table()

caresite_plot_df[, N := N+2]

Specialties_3group <- fread("/data/davis_lab/allie/care_sites/output/Figures/Specialties_NCareSites_Label_ANNOTATED.csv") %>% 
  filter(SpecialtyLong %in% caresite_plot_df$MappedSpecialty) %>% 
  select(MappedSpecialty = SpecialtyLong, Group) %>%
  data.table()

Specialties_3group <- merge(Specialties_3group, unique(caresite_plot_df[,.(MappedSpecialty, N)]), by = "MappedSpecialty")

# # Define colors for each group and alternate light/dark shades
# group_colors <- c("Medical_light" = "#a6cee3", "Medical_dark" = "#1f78b4",
#                   "Surgical_light" = "#b2df8a", "Surgical_dark" = "#33a02c",
#                   "Other_light" = "#fb9a99", "Other_dark" = "#e31a1c")


# Create a base value for each specialty
specialty_base <- Specialties_3group %>%
  mutate(Group = factor(Group, levels = c("Medical", "Surgical", "Other"))) %>%
  arrange(Group, MappedSpecialty) %>%
  distinct(MappedSpecialty, N) %>%
  mutate(base_value = cumsum(N) - N) %>% 
  data.table()

# Order the care sites
caresite_plot_df <- caresite_plot_df  %>% 
  left_join(Specialties_3group %>% select(MappedSpecialty, Group), by=c("MappedSpecialty")) %>%
  group_by(MappedSpecialty) %>%
  ungroup() %>%
  mutate(Group = factor(Group, levels = c("Medical", "Surgical", "Other"))) %>%
  arrange(Group, MappedSpecialty, desc(N)) %>%
  filter(!is.na(Group)) %>% 
  left_join(specialty_base %>% select(MappedSpecialty, base_value), by = c("MappedSpecialty"))

setDT(caresite_plot_df)
setorder(caresite_plot_df, MappedSpecialty, care_site_name)
caresite_plot_df[, id := seq_len(.N), by = MappedSpecialty]
caresite_plot_df[, id := rowid(MappedSpecialty)]
caresite_plot_df[,RowNumber := base_value + id]

caresite_plot_df <- caresite_plot_df %>%
  group_by(MappedSpecialty) %>%
  mutate(Group_Shade = ifelse(cur_group_id() %% 2 == 0, paste(Group, "light", sep = "_"), paste(Group, "dark", sep = "_"))) %>%
  mutate(Specialty_midpoint = median(RowNumber)) %>% 
  ungroup() %>% 
  data.table()

midpoint_labels <- caresite_plot_df %>%
  group_by(MappedSpecialty) %>%
  summarize(midpoint = median(RowNumber)) %>%
  ungroup()

N_spec <- n_distinct(caresite_plot_df$MappedSpecialty)
pal <- rev(rep(brewer.pal(n = 8, name = "Set2"),3)[1:N_spec])

spec_order <- caresite_plot_df %>% arrange(RowNumber) %>% distinct(MappedSpecialty) %>%
  mutate(Color =  pal)

caresite_plot_df <- merge(caresite_plot_df, spec_order, by = "MappedSpecialty")

p <- ggplot(caresite_plot_df, aes(x = RowNumber, y = perc_of_mdd_visits, fill = Color)) +
  geom_bar(stat = "identity", color = "black") +
  geom_label_repel(data = caresite_plot_df %>% filter(perc_of_mdd_visits >= 0.005) %>% arrange(desc(n_mdd_visits)) %>% group_by(MappedSpecialty) %>% slice(1:5), 
                  aes(label = care_site_name),
                  size = 2,
                  box.padding = 0.5, point.padding = 0.5,
                  vjust = 0.5, hjust = 0.5, max.overlaps = 20) +
  scale_x_continuous(breaks = midpoint_labels$midpoint, labels = midpoint_labels$MappedSpecialty) +
  # scale_fill_manual(values = group_colors) +
  scale_fill_identity()  +
  theme_minimal(base_family = "Helvetica") +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 8), 
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"),
        legend.position = "none", 
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) + 
  labs(x = "Care site", y = "Proportion of index MDD visits")

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/earliest_mdd_site_counts_bar.pdf", width = 10, height = 7)
print(p)
dev.off()
```

