---
title: "dep care sites"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table, dplyr, ggplot2, PheWAS, openxlsx, foreach, ggthemr, ggrepel, RColorBrewer, lubridate)
options(scipen=999)

ggthemr_reset()
```

```{r}
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_111524.csv")
care_site_names <- read.xlsx("/data/davis_lab/allie/care_sites/output/CareSite_VisitDetailCount_UPDATED_111524.xlsx") %>%
  select(care_site_id, care_site_name) %>% 
  data.table() %>% 
  unique()

map <- merge(map, care_site_names, by = "care_site_id")
```

```{r}
phecode_map_full <- data.table(unique(rbind(phecode_map_icd10, phecode_map)))
```

## Extract dep ICD codes
```{r load_icd, eval=FALSE}
## load all icd codes from the 2023 SD pull
icd <- readRDS("/data/davis_lab/allie/commonly_used/20230607_sd_pull_x_codes_icd.Rds")
dep_codes <- phecode_map_full[phecode %in% c("296.2", "296.22")]

## restrict to icd codes for dep or depression
icd_dep <- unique(icd[concept_code %in% dep_codes$code, .(GRID, entry_date, entry_datetime, 
                                                          vocabulary_id, concept_code, concept_name, 
                                                          visit_occurrence_id, item_position, age_at_event)])

saveRDS(icd_dep, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression.Rds")
```

```{r}
## load depression/dep icd codes
icd_dep <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression.Rds") %>% 
  select(GRID, entry_date, entry_datetime, vocabulary_id, concept_code, concept_name, visit_occurrence_id, age_at_event) %>% 
  data.table() %>% 
  unique()

icd_dep$entry_datetime <- as_datetime(icd_dep$entry_datetime)

table(icd_dep$vocabulary_id)
```

## Select each individual's *earliest* depression code
```{r}
icd_dep[, earliest_code := min(entry_datetime), by = GRID]
icd_dep_earliest <- unique(icd_dep[entry_datetime == earliest_code, .(GRID,entry_date,entry_datetime,visit_occurrence_id)])

icd_dep_earliest[,visit_ct := .N, by = GRID]
table(icd_dep_earliest$visit_ct)

n_distinct(icd_dep_earliest$GRID) # 203,904 total individuals with dep code
nrow(icd_dep_earliest) # 208,802 total entries
```

### Map depression codes to care sites
Try mapping by visit_occurrence_id AND by date.
```{r, eval=FALSE}
## load all visit data from 2023 SD pull that map to a defined care site (not missing, Unclear, or Administrative)
## (generated in compile_data.R)
vdat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_defined_sites.Rds")
n_distinct(vdat$GRID) # 3,083,596 total individuals with a visit mapped to a "defined" care site 

## get the total individual count per care site and per specialty
vdat[,cs_n_grid := n_distinct(.SD$GRID), by = care_site_id]
vdat[,spec_n_grid := n_distinct(.SD$GRID), by = SpecialtyLong]
cs_grid_counts <- unique(vdat[,.(care_site_id, care_site_name, cs_n_grid, SpecialtyLong)])
spec_grid_counts <- unique(vdat[,.(SpecialtyLong, spec_n_grid)])
setorder(cs_grid_counts, -cs_n_grid)
setorder(spec_grid_counts, -spec_n_grid)
saveRDS(cs_grid_counts, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_defined_sites_individual_counts_by_site.Rds")
saveRDS(spec_grid_counts, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_defined_sites_individual_counts_by_specialty.Rds")

## map earliest depression icd codes to care sites via visit_occurrence_id
icd_dep_visits_by_id <- merge(icd_dep_earliest, vdat, by = c("GRID","visit_occurrence_id"))
saveRDS(icd_dep_visits_by_id, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_visit_id.Rds")

## map earliest depression icd codes to care sites via date
icd_dep_visits_by_date <- merge(icd_dep_earliest, vdat, by = "GRID", allow.cartesian = TRUE)
icd_dep_visits_by_date <- icd_dep_visits_by_date[entry_date >= visit_start_date & entry_date <= visit_end_date]
saveRDS(icd_dep_visits_by_date, "/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_date.Rds")

n_distinct(icd_dep_visits_by_id$GRID) # 139082 individuals with earliest depression code mapping to a visit with a DEFINED CARE SITE by date (68.2% of total with depression)
n_distinct(icd_dep_visits_by_date$GRID) # 150206 individuals with earliest dep code mapping to a visit with a DEFINED CARE SITE by date (73.7% of total with depression)
```
Will map by date 1) because there are more individuals and 2) for consistency with the antidepressant analysis.

```{r}
## load file generated above (joined by date)
icd_dep_visits <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_visit_occurrence_earliest_dep_icd_joined_by_date.Rds")
head(icd_dep_visits)
nrow(icd_dep_visits) # 212406
n_distinct(icd_dep_visits$GRID) # 150206
```

```{r}
## retain unique entries (by GRID, date of ICD code, and care site id)
## there will be duplicates if a care site is mapped to multiple specialties
icd_dep_visits <- unique(icd_dep_visits[,.(GRID,entry_date,care_site_id,care_site_name,SpecialtyLong,cs_n_grid,spec_n_grid)])
nrow(icd_dep_visits) # 186677
n_distinct(icd_dep_visits$GRID) # 150206

nrow(unique(icd_dep_visits[,.(GRID,entry_date,care_site_id)])) # 176580 - number of distinct GRID/care site combos
nrow(unique(icd_dep_visits[,.(GRID,entry_date,care_site_id,care_site_name)])) # 176580 - number of distinct GRID/care site combos (same as above)
nrow(unique(icd_dep_visits[,.(GRID,SpecialtyLong)])) # 181395 - number of distinct GRID/specialty combos
```

Even after restricting to the earliest dep code for each individual and mapping to defined care sites, there are still many individuals with depression codes mapping to multiple different care sites on the same day.
We will keep mutliple entries per individual, but will use the number of unique individuals as the denominator in the care site and specialty frequency calculations.
Otherwise, we would be double-counting individuals with multiple care site visits on the same day. That might bias the results if individuals with dep codes are more likely to have multiple care site visits on the same day, or if certain care sites contribute disproportionately to the total (e.g., radiology or phlebotomy visits, which often co-occur with visits to other sites).

```{r}
paste0("Out of all patients with a depression code, for what % of them does their earliest code map to a visit with a defined specialty? ",
 round((n_distinct(icd_dep_visits$GRID) / n_distinct(icd_dep_earliest$GRID))*100, 2), "%")
```


## Get proportions of individuals with dep codes at each care site and specialty
Normalize by total individual count per care site or specialty.
```{r}
N_dep <- n_distinct(icd_dep_visits$GRID) # 150206 - this will be the denominator for the depression-specific frequencies

icd_dep_visits[,n_dep_per_site := n_distinct(.SD$GRID), by = care_site_id]
icd_dep_visits[,n_dep_per_specialty := n_distinct(.SD$GRID), by = SpecialtyLong]

care_site_summary_dep <- unique(icd_dep_visits[, .(care_site_id, care_site_name, SpecialtyLong, n_dep_per_site, cs_n_grid)])
care_site_summary_dep[,freq_dep_per_site := n_dep_per_site / N_dep]
care_site_summary_dep[,freq_dep_per_site_norm := n_dep_per_site / N_dep / cs_n_grid]

specialty_summary_dep <- unique(icd_dep_visits[, .(SpecialtyLong, n_dep_per_specialty, spec_n_grid)])
specialty_summary_dep[,freq_dep_per_specialty := n_dep_per_specialty / N_dep]
specialty_summary_dep[,freq_dep_per_specialty_norm := n_dep_per_specialty / N_dep / spec_n_grid]
```

## Plots
TODO: create versions of plots using normalized frequencies AND raw frequencies
TODO: modularize plotting functions
```{r}
## restrict to care sites with >=100 depression visits
caresite_plot_df <- care_site_summary_dep %>% 
  select(care_site_id, care_site_name, SpecialtyLong, n_dep_per_site, freq_dep_per_site, freq_dep_per_site_norm) %>%
  filter(n_dep_per_site >=500) %>%
  group_by(SpecialtyLong) %>%
  mutate(N = n()) %>%  ## number of data points per specialty, for plotting purposes
  data.table()

caresite_plot_df[, N := N+2]

## TODO: confirm with John that this is the correct annotation file
Specialties_3group <- fread("/data/davis_lab/allie/care_sites/output/Figures/Descriptives/Specialties_NCareSites_Label_ANNOTATED.csv") %>% 
  filter(SpecialtyLong %in% caresite_plot_df$SpecialtyLong) %>% 
  select(SpecialtyLong, Group) %>%
  data.table()

Specialties_3group <- merge(Specialties_3group, unique(caresite_plot_df[,.(SpecialtyLong, N)]), by = "SpecialtyLong")
Specialties_3group$Group <- factor(Specialties_3group$Group, levels = c("Medical", "Surgical", "Other"))

# Create group to alternate light/dark shades
group_colors <- c("Medical_light" = "#a6cee3", "Medical_dark" = "#1f78b4",
                    "Surgical_light" = "#b2df8a", "Surgical_dark" = "#33a02c",
                    "Other_light" = "#fb9a99", "Other_dark" = "#e31a1c")

# Create a base x-axis value for each specialty
specialty_base <- Specialties_3group
setorder(specialty_base, Group, SpecialtyLong)
specialty_base[, base_value := cumsum(N) - N]

# Order the care sites by specialty and then by care site name
caresite_plot_df <- merge(caresite_plot_df, Specialties_3group[,.(SpecialtyLong, Group)], by = "SpecialtyLong")
caresite_plot_df <- merge(caresite_plot_df, specialty_base[,.(SpecialtyLong, base_value)], by = "SpecialtyLong")
setDT(caresite_plot_df)
setorder(caresite_plot_df, SpecialtyLong, care_site_name)
caresite_plot_df[, id := seq_len(.N), by = SpecialtyLong]
caresite_plot_df[, id := rowid(SpecialtyLong)]
caresite_plot_df[,RowNumber := base_value + id]

caresite_plot_df <- caresite_plot_df %>%
  group_by(SpecialtyLong) %>%
  mutate(Group_Shade = ifelse(cur_group_id() %% 2 == 0, paste(Group, "light", sep = "_"), paste(Group, "dark", sep = "_"))) %>%
  mutate(Specialty_midpoint = median(RowNumber)) %>% 
  ungroup() %>% 
  data.table()

midpoint_labels <- caresite_plot_df %>%
  group_by(SpecialtyLong) %>%
  summarize(midpoint = median(RowNumber)) %>%
  ungroup()

if (FALSE) {
  N_spec <- n_distinct(caresite_plot_df$SpecialtyLong)
  pal <- rev(rep(brewer.pal(n = 8, name = "Set2"),6)[1:N_spec])
  spec_order <- unique(caresite_plot_df[,.(SpecialtyLong)])
  spec_order$Color <- pal
  caresite_plot_df <- merge(caresite_plot_df, spec_order, by = "SpecialtyLong")
}

p <- ggplot(caresite_plot_df, aes(x = RowNumber, y = freq_dep_per_site, fill = Group_Shade)) +
  geom_bar(stat = "identity", color = "black") +
  geom_label_repel(data = caresite_plot_df %>% filter(freq_dep_per_site >= 0.01) %>% arrange(desc(n_dep_per_site)) %>% group_by(SpecialtyLong) %>% slice(1:5), 
                  aes(label = care_site_name),
                  size = 3,
                  box.padding = 0.5, point.padding = 0.5,
                  vjust = 0.5, hjust = 0.5, max.overlaps = 20) +
  scale_x_continuous(breaks = midpoint_labels$midpoint, labels = midpoint_labels$SpecialtyLong) +
  scale_fill_manual(values = group_colors) +
  # scale_fill_identity()  +
  theme_minimal(base_family = "Helvetica") +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 8), 
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"),
        legend.position = "none", 
        panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(x = "Care site", y = "Proportion of index depression diagnoses")

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/earliest_dep_site_counts_bar_mapped_by_date.pdf", width = 10, height = 7)
print(p)
dev.off()
```

```{r}
print(p)
```