---
title: "Care sites MDD application: precision phenotyping"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE, tidy = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width=6, fig.height=4)
pacman::p_load(data.table, dplyr, foreach, PheWAS, ggplot2, plotly, ggthemr, ggpubr, scales, RColorBrewer, ggrepel, kableExtra, broom, openxlsx)
# setwd("/data/davis_lab/allie/care_sites/scripts")
ggthemr_reset()
```

## Define functions
```{r}
#### OR formatting function ####
or_format_func <- function(OR, ci_lower, ci_upper) {
  return(
    paste0(sprintf(
      "%s (%s-%s)",
      formatC(OR, format = "f", digits = 2),
      formatC(ci_lower, format = "f", digits = 2),
      formatC(ci_upper, format = "f", digits = 2)
    ))
  )
}

#### Function to process logistic regression output ####
process_logit <- function(mdl) {
  output <- tidy(mdl)
  output$OR <- exp(output$estimate)
  output$ci.lower <- exp(output$estimate-1.96*output$std.error)
  output$ci.upper <- exp(output$estimate+1.96*output$std.error)
  output
}


#### Rounding function for p-values ####
rounding_func <- function(xx) {
  if (is.na(xx)) {
    return(NA)
  } else if (xx < 1e-3) {
    # y <- formatC(xx, format = "e", digits = 2)
    y <- "<0.001"
  } else if (round(xx,2) <0.01) { # <0.01 and ≥0.001 -> 3 digits 
    y <- formatC(xx, format = "f", digits = 3) 
  } else { # ≥0.01 -> 2 digits
    y <- formatC(xx, format = "f", digits = 2)
  }
  return(as.character(y))
}
```

## Load data, define code-based depression phenotypes
NOTE: the variable names below use "MDD", but the phenotype is actually "depression" (phecode 296.2).
```{r}
## load care site map
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_AML_010425.csv")

## load median age at ICD code
med_age <- fread("/data/davis_lab/shared/phenotype_data/biovu/delivered_data/sd_wide_phewas_covariates/20230607_sd_wide_covariates_121023.txt") %>%
  select(GRID, median_age)

## load covariates and depression PGS
## see compile_data.R for details on how this file was created
dat <- readRDS("/data/davis_lab/allie/care_sites/data/mega_eur_covariates_psych_pgs_010525.Rds") %>%
  filter(genetic_ancestry == "EUR") %>% 
  select(c("GRID", sex = gender_source_value, paste0("PC", 1:10), "pgs.dep.scale")) %>%
  left_join(med_age, by = "GRID") 
```

### Define code-based depression phenotypes
```{r}
## Load ICD code file generated in mdd_care_sites.Rmd, filter to distinct code-dates
# This is everyone in the SD with at least 1 depression code
# Case = has ≥1 code on ≥2 distinct dates
# Control = 0 codes
# Excluded = only 1 code
icd_dep <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression_010525.Rds") %>% 
  select(GRID, entry_date) %>%
  distinct() %>% 
  group_by(GRID) %>% 
  summarise(n = n()) %>%
  mutate(mdd = case_when(n >=2 ~ 1, n == 0 ~ 0, n==1 ~ -9)) %>% 
  select(GRID, mdd, n_dep_code = n)

stopifnot(nrow(icd_dep %>% filter(mdd == 0)) == 0)

## Combine with full cohort
dat <- dat %>% 
  left_join(icd_dep, by = "GRID") %>%
  mutate(mdd = ifelse(is.na(mdd), 0, mdd)) %>%
  mutate(n_dep_code = ifelse(is.na(n_dep_code), 0, n_dep_code))  %>% 
  as.data.frame() %>% data.table()
```

### Remove individuals with missing data
There are 35 individuals in the PGS cohort with missing median age.
This is because these individuals have no ICD codes. The median age was calculated as the median age at the first ICD code, so these individuals have no median age.
```{r}
dat %>% filter(is.na(median_age)) %>% nrow()

dat <- dat %>% filter(!is.na(median_age))
```

```{r}
## load visits
visit <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_mega_grids_visit_occurrence_defined_sites_010525.Rds")
```

## Depression definitions
* Definition 1: standard depression phecode (296.2) on 2+ distinct dates
* Definition 2: 
  * case = has depression phecode (2+ distinct dates) + ≥1 psych site visit (i.e., has depression and has ever seen a psychiatrist)
  * control = no depression phecode
  * excluded = has depression phecode but no psych site visit or only has 1 code
* Definitions 3-5 (not included in violin plots / t-tests as this would get too confusing with the overlapping between cohorts):
  * For N = 3,4,10:
    * case = has depression phecode - N+ distinct dates
    * control = no depression phecode
    * excluded = has only 1-(N-2) distinct depression code-dates
```{r}
psych_sites <- unique(map[SpecialtyLong == "Psychiatry", care_site_id])
psych_grids <- unique(visit[care_site_id %in% psych_sites, GRID])
dat[, psych_cohort := as.numeric(GRID %in% psych_grids)]

dat[, mdd_psych_site := ifelse(mdd==1 & psych_cohort==1, 1,
                                ifelse(mdd==0,0,-9))]

dat[, mdd_pheno := ifelse(mdd_psych_site == 1, "mdd+site", ifelse(mdd == 1, "mdd_no_site", ifelse(mdd == 0, "control", "excluded")))]  
head(dat)                 
```

```{r}
dat %>% 
  group_by(mdd) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

```{r}
dat %>% 
  group_by(mdd_psych_site) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

```{r}
dat %>% 
  group_by(mdd_pheno) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

Create definitions 3-5, for regressions only (not included in violin plots or t-tests).
```{r}
dat[,mdd_3code := ifelse(n_dep_code >= 3, 1, ifelse(n_dep_code == 0, 0, -9))]

dat %>% 
  group_by(mdd_3code) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

```{r}
dat[,mdd_4code := ifelse(n_dep_code >= 4, 1, ifelse(n_dep_code == 0, 0, -9))]

dat %>% 
  group_by(mdd_4code) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

```{r}
dat[,mdd_10code := ifelse(n_dep_code >= 10, 1, ifelse(n_dep_code == 0, 0, -9))]

dat %>% 
  group_by(mdd_10code) %>%
  summarise(n = n(), min_n_dep_code = min(n_dep_code), max_n_dep_code = max(n_dep_code))
```

## Raw PGS comparison
```{r}
my_comparisons <- list( c("Control", "2+ Codes Only"), c("Control", "2+ Codes + Psych. Visit"), c("2+ Codes Only", "2+ Codes + Psych. Visit") )

dat_plot <-  dat %>% 
  filter(mdd_pheno != "excluded") %>%
  mutate(mdd_pheno = factor(mdd_pheno, levels = c("control", "mdd_no_site", "mdd+site"), labels = c("Control", "2+ Codes Only", "2+ Codes + Psych. Visit"))) %>% 
  mutate(mdd_pheno = as.character(mdd_pheno)) %>%
  data.table()

p1 <- dat_plot %>%
  ggplot(aes(x = mdd_pheno, fill = mdd_pheno, pgs.dep.scale)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.3) +
  scale_y_continuous(expand = expansion(mult = 0.05)) +
  labs(x = "Phenotype", y = "Scaled Depression Polygenic Score", fill = "") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif") +
  theme_bw() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 7, angle = 20, hjust = 1), 
        axis.title.x = element_text(margin = margin(t = 1)))

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_raw_pgs_comparison_010525.pdf", width = 6, height = 5)
print(p1)
dev.off()
```

```{r}
print(p1)
```

### Stats
```{r}
## Codes only vs Controls
t.test(dat_plot[ mdd_pheno == "2+ Codes Only", pgs.dep.scale], dat_plot[ mdd_pheno == "Control", pgs.dep.scale])

## Codes+Psych vs Controls
t.test(dat_plot[ mdd_pheno == "2+ Codes + Psych. Visit", pgs.dep.scale], dat_plot[ mdd_pheno == "Control", pgs.dep.scale])

## Codes+Psych vs Codes only
t.test(dat_plot[ mdd_pheno == "2+ Codes + Psych. Visit", pgs.dep.scale], dat_plot[ mdd_pheno == "2+ Codes Only", pgs.dep.scale])
```

## PGS regression analysis
```{r}
## outcome = depression phecode (2+ codes)
mdl_mdd <- glm(mdd ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd!=-9], family = "binomial")

res_mdd <- process_logit(mdl_mdd) %>% 
  mutate(pheno = "2+ Codes", 
         n_total = nrow(dat[mdd!=-9]), 
         n_cases = nrow(dat[mdd==1]), 
         n_controls = nrow(dat[mdd==0]))

## outcome = depression phecode (3+ codes)
mdl_mdd_3code <- glm(mdd_3code ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd_3code!=-9], family = "binomial")

res_mdd_3code <- process_logit(mdl_mdd_3code) %>% 
  mutate(pheno = "3+ Codes", 
         n_total = nrow(dat[mdd_3code!=-9]), 
         n_cases = nrow(dat[mdd_3code==1]), 
         n_controls = nrow(dat[mdd_3code==0]))

## outcome = depression phecode (4+ codes)
mdl_mdd_4code <- glm(mdd_4code ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd_4code!=-9], family = "binomial")

res_mdd_4code <- process_logit(mdl_mdd_4code) %>% 
  mutate(pheno = "4+ Codes", 
         n_total = nrow(dat[mdd_4code!=-9]), 
         n_cases = nrow(dat[mdd_4code==1]), 
         n_controls = nrow(dat[mdd_4code==0]))

## outcome = depression phecode (10+ codes)
mdl_mdd_10code <- glm(mdd_10code ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd_10code!=-9], family = "binomial")

res_mdd_10code <- process_logit(mdl_mdd_10code) %>% 
  mutate(pheno = "10+ Codes", 
         n_total = nrow(dat[mdd_10code!=-9]), 
         n_cases = nrow(dat[mdd_10code==1]), 
         n_controls = nrow(dat[mdd_10code==0]))

## outcome = depression phecode (2+ codes) + psych site visit
mdl_mdd_psych_site <- glm(mdd_psych_site ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd_psych_site!=-9], family = "binomial")

res_mdd_psych_site <- process_logit(mdl_mdd_psych_site) %>% 
  mutate(pheno = "2+ Codes +\nPsych. Visit", 
         n_total = nrow(dat[mdd_psych_site!=-9]), 
         n_cases = nrow(dat[mdd_psych_site==1]),
         n_controls = nrow(dat[mdd_psych_site==0]))

## combine results
res_combined <- data.table(rbind(res_mdd, res_mdd_psych_site, res_mdd_3code, res_mdd_4code, res_mdd_10code))

res_combined[, OR_format := apply(.SD, 1, function(x) or_format_func(x[1], x[2], x[3])), .SDcols = c("OR", "ci.lower", "ci.upper")]
res_combined$p.value.fmt <- sapply(res_combined$p.value, rounding_func)
res_combined$pheno <- factor(res_combined$pheno, levels = rev(c("2+ Codes", "3+ Codes", "4+ Codes", "10+ Codes", "2+ Codes +\nPsych. Visit")))
res_combined[, pheno_color := as.character(pheno == "2+ Codes +\nPsych. Visit")]

res_combined %>% 
  filter(term == "pgs.dep.scale") %>% 
  select(pheno, OR_format, p.value.fmt, n_cases, n_controls) %>%
  arrange(pheno)
```

### Forest plot
```{r}
p2 <- res_combined %>% 
  filter(term == "pgs.dep.scale") %>% 
  ggplot(aes(x = pheno, y = OR, ymin = ci.lower, ymax = ci.upper, color = pheno_color)) +
  geom_point(size = 3) +
  geom_errorbar(width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(expand = expansion(mult = 0.05))  +
  scale_color_manual(values = c("TRUE" = "#E64B35")) + 
  labs(x = "Phenotype", y = "Odds ratio (95% CI)") +
  coord_flip() +
  theme_bw() +
  guides(color = "none") +
  theme(axis.text.y = element_text(size = 8), 
        axis.title.y = element_text(margin = margin(r = 0)),
        axis.title.x = element_text(margin = margin(t = 5)),
        panel.grid.major.y = element_blank(),
        plot.margin =  margin(t = 0.2, r = 0.2, b = 0.2, l = 0.1, unit = "cm")) 

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_regression_forest_plot_010525.pdf", width = 6, height = 4)
print(p2)
dev.off()
```

```{r}
print(p2)
```

### Combined plot
```{r}
p3 <- ggarrange(p1, p2, ncol = 2, nrow = 1, widths = c(1, 0.85), labels = c("A", "B"))

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_regression_violins_forest_combined_010525.pdf", width = 7, height = 5)
print(p3)
dev.off()
```

```{r}
print(p3)
```

## Supplemental tables
```{r}
supp_table <- copy(res_combined) %>% 
  select(term, beta = estimate, std.error, statistic, OR, ci.lower, ci.upper, p.value, n_total, n_cases, pheno) %>%
  as.data.frame() %>% data.table()

supp_table[pheno == "2+ Codes +\nPsych. Visit", pheno := "2+ Codes + Psych. Visit"]
supp_table[, pheno := as.character(pheno)]
setnames(supp_table, "pheno", "Phenotype")

## write to excel
wb <- createWorkbook()
for (pheno in unique(supp_table$Phenotype)) {
  addWorksheet(wb, pheno)
  writeData(wb, pheno, supp_table[Phenotype == pheno] %>% select(-Phenotype))
}
saveWorkbook(wb, paste0("/data/davis_lab/allie/care_sites/output/mdd_application/SupplementalTable_depression_pgs_regression_results_", format(Sys.Date(), "%m%d%y"), ".xlsx"), overwrite = TRUE)
```

### Save analysis cohort for supplemental table created in depression_analyses_descriptive_table.R
```{r}
dat %>% filter(mdd!=-9) %>%
  select(GRID, sex, median_age, mdd, mdd_pheno, psych_cohort, mdd_psych_site) %>%
  fwrite("/data/davis_lab/allie/care_sites/data/mdd_application/mdd_pgs_regression_cohort_010525.csv")
```