---
title: "Care sites MDD application: precision phenotyping"
author: "Allie Lake"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE, tidy = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width=6, fig.height=4)
pacman::p_load(data.table, dplyr, foreach, PheWAS, ggplot2, plotly, ggthemr, ggpubr, scales, RColorBrewer, ggrepel, kableExtra, broom, openxlsx)
# setwd("/data/davis_lab/allie/care_sites/scripts")
ggthemr_reset()
```

## Define functions
```{r}
#### OR formatting function ####
or_format_func <- function(OR, ci_lower, ci_upper) {
  return(
    paste0(sprintf(
      "%s (%s-%s)",
      formatC(OR, format = "f", digits = 2),
      formatC(ci_lower, format = "f", digits = 2),
      formatC(ci_upper, format = "f", digits = 2)
    ))
  )
}

#### Function to process logistic regression output ####
process_logit <- function(mdl) {
  output <- tidy(mdl)
  output$OR <- exp(output$estimate)
  output$ci.lower <- exp(output$estimate-1.96*output$std.error)
  output$ci.upper <- exp(output$estimate+1.96*output$std.error)
  output
}


#### Rounding function for p-values ####
rounding_func <- function(xx) {
  if (is.na(xx)) {
    return(NA)
  } else if (xx < 1e-3) {
    # y <- formatC(xx, format = "e", digits = 2)
    y <- "<0.001"
  } else if (round(xx,2) <0.01) { # <0.01 and ≥0.001 -> 3 digits 
    y <- formatC(xx, format = "f", digits = 3) 
  } else { # ≥0.01 -> 2 digits
    y <- formatC(xx, format = "f", digits = 2)
  }
  return(as.character(y))
}
```

## Load data, define code-based depression phenotypes
NOTE: the variable names below use "MDD", but the phenotype is actually "depression" (phecode 296.2).
```{r}
## load care site map
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_AML_010425.csv")

## load median age at ICD code
med_age <- fread("/data/davis_lab/shared/phenotype_data/biovu/delivered_data/sd_wide_phewas_covariates/20230607_sd_wide_covariates_121023.txt") %>%
  select(GRID, median_age)

## load covariates and depression PGS
## see compile_data.R for details on how this file was created
dat <- readRDS("/data/davis_lab/allie/care_sites/data/mega_eur_covariates_psych_pgs_010525.Rds") %>%
  filter(genetic_ancestry == "EUR") %>% 
  select(c("GRID", sex = gender_source_value, paste0("PC", 1:10), "pgs.dep.scale")) %>%
  left_join(med_age, by = "GRID") 
```

### Define code-based depression phenotypes
```{r}
## Load ICD code file generated in mdd_care_sites.Rmd, filter to distinct code-dates
# This is everyone in the SD with at least 1 depression code
# Case = has ≥1 code on ≥2 distinct dates
# Control = 0 codes
# Excluded = only 1 code
icd_dep <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_x_codes_icd_depression_010525.Rds") %>% 
  select(GRID, entry_date) %>%
  distinct() %>% 
  group_by(GRID) %>% 
  summarise(n = n()) %>%
  mutate(mdd = case_when(n >=2 ~ 1, n == 0 ~ 0, n==1 ~ -9)) %>% 
  select(GRID, mdd)

stopifnot(nrow(icd_dep %>% filter(mdd == 0)) == 0)

## Combine with full cohort
dat <- dat %>% 
  left_join(icd_dep, by = "GRID") %>%
  mutate(mdd = ifelse(is.na(mdd), 0, mdd)) %>%
  as.data.frame() %>% data.table()
```

### Remove individuals with missing data
There are 35 individuals in the PGS cohort with missing median age.
This is because these individuals have no ICD codes. The median age was calculated as the median age at the first ICD code, so these individuals have no median age.
```{r}
dat %>% filter(is.na(median_age)) %>% nrow()

dat <- dat %>% filter(!is.na(median_age))
```

```{r}
## load visits
visit <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/20230607_sd_pull_mega_grids_visit_occurrence_defined_sites_010525.Rds")
```

## Depression definitions
* Definition 1: standard depression phecode (296.2)
* Definition 2: 
  * case = has depression phecode (296.2) + ≥1 psych site visit (i.e., has depression and has ever seen a psychiatrist)
  * control = no depression phecode (296.2)
  * excluded = has depression phecode (296.2) but no psych site visit or is N/A for depression phecode
```{r}
psych_sites <- unique(map[SpecialtyLong == "Psychiatry", care_site_id])
psych_grids <- unique(visit[care_site_id %in% psych_sites, GRID])
dat[, psych_cohort := as.numeric(GRID %in% psych_grids)]

dat[, mdd_psych_site := ifelse(mdd==1 & psych_cohort==1, 1,
                                ifelse(mdd==0,0,-9))]

dat[, mdd_pheno := ifelse(mdd_psych_site == 1, "mdd+site", ifelse(mdd == 1, "mdd_no_site", ifelse(mdd == 0, "control", "excluded")))]  
head(dat)                 
```

```{r}
dat %>% 
  group_by(mdd) %>%
  summarise(n = n())
```

```{r}
dat %>% 
  group_by(mdd_psych_site) %>%
  summarise(n = n())
```

```{r}
dat %>% 
  group_by(mdd_pheno) %>%
  summarise(n = n())
```

## Raw PGS comparison
```{r}
my_comparisons <- list( c("Control", "Code only"), c("Control", "Code + psych. visit"), c("Code only", "Code + psych. visit") )

dat_plot <-  dat %>% 
  filter(mdd_pheno != "excluded") %>%
  mutate(mdd_pheno = factor(mdd_pheno, levels = c("control", "mdd_no_site", "mdd+site"), labels = c("Control", "Code only", "Code + psych. visit"))) %>% 
  mutate(mdd_pheno = as.character(mdd_pheno)) %>%
  data.table()

p1 <- dat_plot %>%
  ggplot(aes(x = mdd_pheno, fill = mdd_pheno, pgs.dep.scale)) +
  geom_violin() +
  geom_boxplot(width = 0.3) +
  labs(x = "Phenotype", y = "Scaled depression PGS", fill = "") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test", label = "p.signif") +
  theme_bw() +
  theme(legend.position = "none")

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_raw_pgs_comparison_010525.pdf", width = 6, height = 5)
print(p1)
dev.off()
```

```{r}
print(p1)
```

## PGS regression analysis
```{r}
mdl_mdd <- glm(mdd ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd!=-9], family = "binomial")

res_mdd <- process_logit(mdl_mdd) %>% 
  mutate(pheno = "MDD", 
         n_total = nrow(dat[mdd!=-9]), 
         n_cases = nrow(dat[mdd==1]), 
         n_controls = nrow(dat[mdd==0]))

mdl_mdd_psych_site <- glm(mdd_psych_site ~ pgs.dep.scale + sex + median_age + PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, 
           data = dat[mdd_psych_site!=-9], family = "binomial")

res_mdd_psych_site <- process_logit(mdl_mdd_psych_site) %>% 
  mutate(pheno = "MDD+psych_site", 
         n_total = nrow(dat[mdd_psych_site!=-9]), 
         n_cases = nrow(dat[mdd_psych_site==1]),
         n_controls = nrow(dat[mdd_psych_site==0]))

res_combined <- data.table(rbind(res_mdd, res_mdd_psych_site))

res_combined[, OR_format := apply(.SD, 1, function(x) or_format_func(x[1], x[2], x[3])), .SDcols = c("OR", "ci.lower", "ci.upper")]
res_combined$p.value.fmt <- sapply(res_combined$p.value, rounding_func)

res_combined %>% 
  filter(term == "pgs.dep.scale") %>% 
  select(pheno, OR_format, p.value.fmt) %>%
  kable() %>%
  kable_styling()
```

### Forest plot
```{r}
p2 <- res_combined %>% 
  filter(term == "pgs.dep.scale") %>% 
  mutate(pheno = factor(pheno, levels = c("MDD", "MDD+psych_site"), labels = c("Code only", "Code + psych. visit"))) %>%
  ggplot(aes(x = pheno, y = OR, ymin = ci.lower, ymax = ci.upper)) +
  geom_point(size = 3) +
  geom_errorbar(width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_discrete(expand = c(0, 0.4))  +
  labs(x = "Phenotype", y = "Odds ratio (95% CI)") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.y = element_text(size = 8), 
        axis.title.y = element_text(margin = margin(r = 10)),
        axis.title.x = element_text(margin = margin(t = 10)),
        panel.grid.major.y = element_blank()) 

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_regression_forest_plot_010525.pdf", width = 3, height = 1.5)
print(p2)
dev.off()
```

```{r}
print(p2)
```

### Combined plot
```{r}
p3 <- ggarrange(p1, p2, ncol = 2, nrow = 1, widths = c(1, 1))

pdf("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_regression_violins_forest_combined_010525.pdf", width = 6, height = 3)
print(p3)
dev.off()
```

```{r}
print(p3)
```

## Supplemental tables
```{r}
supp_table <- copy(res_combined) %>% 
  select(term, beta = estimate, std.error, statistic, OR, ci.lower, ci.upper, p.value, n_total, n_cases, pheno) %>%
  data.table()

## write to excel
wb <- createWorkbook()
addWorksheet(wb, "Code only")
writeData(wb, "Code only", supp_table[pheno == "MDD"] %>% select(-pheno))
addWorksheet(wb, "Code + psych. visit")
writeData(wb, "Code + psych. visit", supp_table[pheno == "MDD+psych_site"] %>% select(-pheno))
saveWorkbook(wb, "/data/davis_lab/allie/care_sites/output/mdd_application/SupplementalTable_depression_pgs_regression_results_010525.xlsx", overwrite = TRUE)
```

### Save analysis cohort for supplemental table created in depression_analyses_descriptive_table.R
```{r}
dat %>% filter(mdd!=-9) %>%
  select(GRID, sex, median_age, mdd, mdd_pheno, psych_cohort, mdd_psych_site) %>%
  fwrite("/data/davis_lab/allie/care_sites/data/mdd_application/mdd_pgs_regression_cohort_010525.csv")
```