---
title: "Care sites application #2a: stratified depression PGS PheWAS"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

## setup and load data
```{r setup, include = FALSE, message = FALSE, warning = FALSE, tidy = TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width=6, fig.height=4)
pacman::p_load(data.table, dplyr, foreach, PheWAS, ggplot2, plotly, ggpubr, scales, RColorBrewer, ggrepel)
setwd("/data/davis_lab/allie/care_sites/scripts")
```

### Load data compiled in compile_data.R
```{r}
## load care site map
map <- fread("/data/davis_lab/allie/care_sites/output/CareSiteMap_Multispecialty_Long_UPDATED_092424.csv")

## load covariates, PGS, and phecodes
dat <- readRDS("/data/davis_lab/allie/care_sites/data/mdd_application/mega_eur_afr_covariates_pgs_phecodes_092424.Rds")
setnames(dat, "pgs.dep.scale", "pgs.mdd.scale")
dat <- dat[genetic_ancestry == "EUR"] # subset to EUR ancestry for now

## visit info
visit <- fread("/data/davis_lab/allie/care_sites/data/mdd_application/mega_grids_visits_cleaned_dates_non_null_sites.csv")
```

## construct primary care cohort and downsampled full cohort
```{r}
pc_sites <- unique(map[SpecialtyLong == "PrimaryCare", care_site_id])
pc_grids <- unique(visit[care_site_id %in% pc_sites, GRID])
dat[, pc_cohort := GRID %in% pc_grids]
```

### downsample full cohort to match primary care cohort
```{r}
n_downsample <- nrow(dat[pc_cohort == TRUE & genetic_ancestry == "EUR"])

set.seed(123)
grids_downsample <- sample(unique(dat$GRID), n_downsample) 
dat[, downsample := GRID %in% grids_downsample]
```

```{r}
table(dat$pc_cohort)
table(dat$downsample)
```

## run phewas (EUR-ancestry cohort for now)
```{r}
phe_cols<- names(dat)[grepl("^[[:digit:]]", names(dat))]

## phecode case counts
# count number of cases per sex per phecode
case_counts <- function(Sex) {
  case_counts <- data.frame(count = unlist(lapply(dat[sex == Sex, phe_cols, with = F], function(phe) sum(phe, na.rm = T))))
  case_counts$phecode <- rownames(case_counts)
  case_counts <- data.table(case_counts[, c(2, 1)])
  case_counts
}

counts_f <- case_counts("F")
counts_m <- case_counts("M")
counts_both <- merge(counts_f, counts_m, by = "phecode", suffixes = c(".f", ".m"))
counts_both <- addPhecodeInfo(counts_both)
counts_both <- counts_both[, .(phecode, description, group, count.f, count.m)]
counts_both[, count.tot := count.f + count.m]

## remove phecodes with <50 cases overall or <5 cases in one sex
phe.keep <- counts_both[count.tot >= 50 & count.f >= 10 & count.m >= 10, phecode]

## phewas function
my_phewas <- function(dt, pred_col = "pgs.mdd.scale", covar = c("sex", "median_age", paste0("PC", 1:10))) {

  message("Running PheWAS")
  message("Predictor: ", paste(pred_col))
  message("Covariates: ", paste(covar, collapse = ", "))

  res <- phewas(
      data = dt, phenotypes = phe.keep[!(phe.keep %in% covar)], predictors = pred_col,
      covariates = covar, additive.genotypes = F, min.records = 50,
      significance.threshold = "bonferroni"
  )

  res <- data.table(addPhecodeInfo(res, groupcolors = T, groups = T, groupnums = T))
  setorder(res, p, -OR)
  res[, ci.lower := exp(beta - 1.96 * SE)]
  res[, ci.upper := exp(beta + 1.96 * SE)]

  res <- res[!is.na(beta)]

  res <- res %>% 
    select(-predictor, -adjustment, -type, -HWE_p, -allele_freq, -n_no_snp) %>%
    data.table()

  return(res)
}
```

```{r, eval = FALSE}
res_phewas_full <- my_phewas(dat)
saveRDS(res_phewas_full, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_111824.Rds")

res_phewas_full_downsample <- my_phewas(dat[downsample == TRUE])
saveRDS(res_phewas_full_downsample, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_downsample_111824.Rds")

res_phewas_pc <- my_phewas(dat[pc_cohort == TRUE])
saveRDS(res_phewas_pc, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_pc_cohort_111824.Rds")
```

### phewas controlling for depression phecode
```{r, eval = FALSE}
res_phewas_full <- my_phewas(dat, covar = c("sex", "median_age","296.2", paste0("PC", 1:10)))
saveRDS(res_phewas_full, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_111824.Rds")

res_phewas_full_downsample <- my_phewas(dat[downsample == TRUE], covar = c("sex", "median_age","296.2", paste0("PC", 1:10)))
saveRDS(res_phewas_full_downsample, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_downsample_adj_mdd_111824.Rds")

res_phewas_pc <- my_phewas(dat[pc_cohort == TRUE], covar = c("sex", "median_age","296.2", paste0("PC", 1:10)))
saveRDS(res_phewas_pc, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_pc_cohort_adj_mdd_111824.Rds")
```

```{r}
res_phewas_full <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_111824.Rds")
res_phewas_full_downsample <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_downsample_111824.Rds")
res_phewas_pc <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_pc_cohort_111824.Rds")

res_phewas_full_adj_mdd <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_111824.Rds")
res_phewas_full_downsample_adj_mdd <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_downsample_adj_mdd_111824.Rds")
res_phewas_pc_adj_mdd <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_pc_cohort_adj_mdd_111824.Rds")
```

## phewas comparisons
```{r}
# create my own color: fixed mapping between group and color
group_list <- c("circulatory system", "congenital anomalies", "dermatologic", "digestive", "endocrine/metabolic", "genitourinary", "hematopoietic", "infectious diseases", "injuries & poisonings", "mental disorders", "musculoskeletal", "neoplasms", "neurological", "pregnancy complications", "respiratory", "sense organs", "symptoms")
color_groups <- data.frame(group = group_list, color = rep(brewer.pal(8, "Set2"), 3)[1:length(group_list)])

my_scatter <- function(dt) {
  r2 <- cor(dt$beta.x, dt$beta.y)^2

  dt[, bonf_only_y := ifelse(bonferroni.x==FALSE & bonferroni.y==TRUE, TRUE, FALSE)]
  dt[, bonf_only_x := ifelse(bonferroni.y==FALSE & bonferroni.x==TRUE, TRUE, FALSE)]

  ggplot(dt, aes(x = beta.x, y = beta.y, color = group, fill = group, label = description)) +
    geom_point() +
    geom_point(data = dt[phenotype %in% c("071", "071.1")], color = "black", size = 3) +
    geom_text(data = dt[phenotype %in% c("071", "071.1")], color = "black", size = 2) +
    scale_color_manual(values = color_groups$color, breaks = color_groups$group) +
    scale_fill_manual(values = color_groups$color, breaks = color_groups$group) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    annotate("text", x = 0, y = 0.5, label = paste("r^2 =", round(r2, 2)), size = 3) +
    theme_minimal() +
    guides(fill="none")
}
```

### full cohort vs primary care
```{r}
full_downsample_vs_pc <- merge(res_phewas_full_downsample, res_phewas_pc, 
                    by = c("phenotype","description","groupnum","group","color"), 
                    suffixes = c(".x", ".y"))

p <- my_scatter(full_downsample_vs_pc) +
  labs(title = "Full cohort vs primary care cohort", x = "Full cohort", y = "Primary care cohort") 

ggplotly(p)
```

### full cohort vs primary care - adjusting for psychosis
```{r}
full_vs_pc_adj_mdd <- merge(res_phewas_full_downsample_adj_mdd, res_phewas_pc_adj_mdd, 
                    by = c("phenotype","description","groupnum","group","color"), 
                    suffixes = c(".x", ".y"))

p <- my_scatter(full_vs_pc_adj_mdd) +
  labs(x = "Full cohort", y = "Primary care cohort") 

ggplotly(p)
```

## phewas controlling for CCC site
```{r}
ccc_grids <- unique(visit[care_site_id == 5023, GRID])
dat[, ccc_cohort := GRID %in% ccc_grids]
table(dat$ccc_cohort)
```

### run adjusted and stratified phewas
```{r, eval = FALSE}
res_phewas_full_adj_mdd_ccc  <- my_phewas(dat, covar = c("sex", "median_age","296.2", "ccc_cohort", paste0("PC", 1:10)))
saveRDS(res_phewas_full_adj_mdd_ccc, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_ccc_111824.Rds")

res_phewas_full_adj_mdd_rm_ccc  <- my_phewas(dat[ccc_cohort==FALSE], covar = c("sex", "median_age","296.2", paste0("PC", 1:10)))
saveRDS(res_phewas_full_adj_mdd_rm_ccc, "/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_rm_ccc_111824.Rds")
```

```{r}
res_phewas_full_adj_mdd_ccc <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_ccc_111824.Rds")
res_phewas_full_adj_mdd_rm_ccc <- readRDS("/data/davis_lab/allie/care_sites/output/mdd_application/mdd_pgs_phewas_full_cohort_adj_mdd_rm_ccc_111824.Rds")
```

### comparison
```{r}
adj_ccc_comparison <- merge(res_phewas_full_adj_mdd, res_phewas_full_adj_mdd_ccc, 
                    by = c("phenotype","description","groupnum","group","color"), 
                    suffixes = c(".x", ".y"))

p <- my_scatter(adj_ccc_comparison) +
  labs(title = "", x = "Full cohort", y = "Full cohort, adjusting for CCC visit") 

ggplotly(p)
```

```{r}
rm_ccc_comparison <- merge(res_phewas_full_adj_mdd, res_phewas_full_adj_mdd_rm_ccc, 
                    by = c("phenotype","description","groupnum","group","color"), 
                    suffixes = c(".x", ".y"))

p <- my_scatter(rm_ccc_comparison) +
  labs(title = "", x = "Full cohort", y = "Full cohort, removing those with a CCC visit") 

ggplotly(p)
```